\pkg{depmixS4} : An \proglang{R}-package for hidden Markov models

### 2009-07-14 papers/jss/article.tex
minor stuff, typo's and the like
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/depmix@303 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

	\pkg{depmixS4} implements a general framework for defining and estimating
	 dependent mixture models in the \proglang{R} programming language
	\citep{R2009}. This includes standard Markov models, latent/hidden Markov
	models, and latent class and finite mixture distribution models. The models can
	be fitted on mixed multivariate data with distributions from the \code{glm} 
	family, the logistic multinomial, or the multivariate normal distribution. Other 
	distributions can be added easily, and an example is provided. Parameter 
	estimation is done through an EM algorithm or by a direct optimization approach using
	the\pkg{Rdonlp2} optimization routine when contraints are imposed on the parameters. 
	Parameters can be estimated subject to general linear constraints.  
}

@@

The latent Markov model is commonly associated with data of this type,
although usually only multinomial response variables are considered.  However,
common estimation procedures, such as those implemented in
\citet{Pol1996} are not suitable for long time series due to underflow
problems.  In contrast, the hidden Markov model is typically only used
for `long' univariate time series 
\cite[see e.g.][, chapter~1 for an overview of examples]{Cappe2005}.  
In the next sections, the
likelihood and estimation procedures for the dependent mixture model is
described for data of the above form.  These models are called
dependent mixture models because one of the authors (Ingmar Visser)
thought it was time for a new name for these models\footnote{Only
later did I find out that \citet{Leroux1992} already coined the term
dependent mixture models in an application with hidden Markov mixtures
of Poisson count data.}.

>>>>

	\pkg{depmixS4} implements a general framework for defining and estimating
	 dependent mixture models in the \proglang{R} programming language
	\citep{R2009}. This includes standard Markov models, latent/hidden Markov
	models, and latent class and finite mixture distribution models. The models can
	be fitted on mixed multivariate data with distributions from the \code{glm} 
	family, the logistic multinomial, or the multivariate normal distribution. Other 
	distributions can be added easily, and an example is provided. Parameter 
	estimation is done through an EM algorithm or by a direct optimization approach using
	the \pkg{Rdonlp2} optimization routine when contraints are imposed on the parameters. 
	Parameters can be estimated subject to general linear constraints.  
}

@@

The latent Markov model is commonly associated with data of this type,
although usually only multinomial response variables are considered.  However,
common estimation procedures, such as those implemented in
\citet{Pol1996} are not suitable for long time series due to underflow
problems.  In contrast, the hidden Markov model is typically only used
for `long' univariate time series 
\citep[][, chapter~1]{Cappe2005}.  
In the next sections, the
likelihood and estimation procedures for the dependent mixture model is
described for data of the above form.  These models are called
dependent mixture models because one of the authors (Ingmar Visser)
thought it was time for a new name for these models\footnote{Only
later did I find out that \citet{Leroux1992} already coined the term
dependent mixture models in an application with hidden Markov mixtures
of Poisson count data.}.
